---
title: "descriptiv analysis of the penguin dataset"
author: "Vanessa"
format: 
  docx:
    reference-doc: C:/Users/Vanessa Fleischer/template_quarto.docx
    toc: true #table of content
editor: visual
execute: 
  echo: false # jammel needs lower letters -->  the underling source code of a function won´t be in the document, just the output
  warning: false # no warnings in the report
  output: asis # output as it is, no formating
#figure setting
fig-dpi: 300 # graphs to publish 600 
fig-width: 6 # 6 and 8 fill a page with the figure
fig-height: 8 # 3.2 = 2 figures on one page
tbl-cap-location: top #Table caption top and not underneath
---

# descriptiv analysis with summarize / across

```{r}
pacman::p_load(conflicted,tidyverse,wrappedtools,
               flextable)

conflicts_prefer(dplyr::filter)

data(penguins)

clean_data_penguins <- penguins |>
  drop_na()
```

<br>

## 1 variable / 1 function / no groups 

<br>

```{r}
summarize(.data = clean_data_penguins,
          MeanSD=meansd(body_mass)) |> 
  flextable()
```

<br>

## 1 variable / 1 function / subgroups

### groups in rows

<br>

```{r}
clean_data_penguins |> 
  group_by(species) |> 
  summarize(MeanSD=meansd(body_mass),
          .groups = 'drop') |> 
  flextable()|>
  set_table_properties(width=0.4, layout="autofit")

```

###  groups in columns

<br>

```{r}
clean_data_penguins |> 
  group_by(species) |> 
summarize(MeanSD=meansd(body_mass),
          .groups = 'drop') |>
  pivot_wider(names_from = species, # create new columnes!
              values_from = MeanSD,
              names_glue = "{species}") |>#optional
  flextable()|> 
  separate_header(split = ": ")|>
  set_table_properties(width=0.4, layout="autofit")
```

<br>

```{r}
clean_data_penguins |> 
  group_by(species) |> 
summarize(MeanSD=meansd(body_mass),
          .groups = 'drop') |>
  pivot_longer(
    cols = "MeanSD",
    names_to = "  ",
    values_to = "MeanSD")|>
  pivot_wider(names_from = species, # create new columnes!
              values_from = MeanSD) |>#optional
  flextable()|> 
  separate_header(split = ": ")
```

<br>

## 1 variable / 2 functions / no groups 

<br>

```{r}
summarize(.data = clean_data_penguins,
          `MeanSD body_mass`=meansd(body_mass),
          `MedianQuartiles body_mass`=median_quart(body_mass)) |> 
  flextable()|>
  set_table_properties(width=0.4, layout="autofit")
```

<br>

## 1 variable / 2 functions / subgroups

### groups in rows

<br>

```{r}
clean_data_penguins |> 
  group_by(species) |> 
  summarize( `MeanSD body_mass`=meansd(body_mass, roundDig = 3),
          `MedianQuartiles body_mass`=median_quart(body_mass, roundDig = 3),
          .groups = 'drop') |> 
  flextable() |>
  separate_header(split = " ") |>
  set_table_properties(width=0.6, layout="autofit")
```

<br>

###  groups in columns 

<br>

```{r}
clean_data_penguins |> 
  group_by(species) |> 
  summarize(`MeanSD body_mass`=meansd(body_mass),
          `MedianQuartiles body_mass`=median_quart(body_mass),
          .groups = 'drop') |> 
  pivot_longer(
    cols = starts_with("M"),
    names_to = "Statistics",
    values_to = "estimate") |>
  pivot_wider(names_from = species,
              values_from = estimate) |>
  flextable()|>
  separate_header(split = ": ")|>
  set_table_properties(width=0.8, layout="autofit")

```

<br>

## all variables / 1 function / no groups 

<br>

```{r}

numvars <-
  ColSeeker(
    data = penguins, # can be omitted, as it is the default
    namepattern = "_"
  )
```

<br>

###  no function arguments 

<br>

```{r}
clean_data_penguins |> 
  summarize(across(all_of(numvars$names),
                   .fns=~meansd(.x))) |> 
  flextable()
```

<br>

###  with function arguments 

<br>

```{r}
clean_data_penguins |> 
  summarize(across(all_of(numvars$names),
                   .fns=~meansd(.x,add_n = TRUE))) |> 
  flextable()|> 
  set_table_properties(width=.7, layout="autofit")
```

<br>

### Variables in rows

<br>

```{r}
clean_data_penguins |> 
  summarize(across(all_of(numvars$names),
                   .fns=~meansd(.x,add_n = TRUE))) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "MeanSD") |>
  flextable() |> 
  set_table_properties(width=.7, layout="autofit")
```

<br>

## all variables / 1 function / subgroups

### group in rows

<br>

```{r}
clean_data_penguins |> 
  group_by(species) |> 
  summarize(across(all_of(numvars$names),
                   ~meansd(.x,add_n = TRUE)),
            .groups = 'drop') |> 
  flextable()|> 
  set_table_properties(width=.7, layout="autofit")
```

<br>

### Variables in rows

<br>

```{r}
clean_data_penguins |> 
  group_by(species) |> 
  summarize(across(all_of(numvars$names),
                   ~meansd(.x,add_n = TRUE)),
            .groups = 'drop') |> 
  pivot_longer(
    cols = -species,
    names_to = "Variable",
    values_to = "MeanSD") |>
  pivot_wider(names_from = species,
              values_from = MeanSD,
              names_prefix = "species: ") |>
  flextable() |>
  separate_header(split = ": ")|> 
  set_table_properties(width=.7, layout="autofit")
```

<br>

## all variables / 2 function / no groups 

<br>

```{r}

clean_data_penguins |> 
  summarize(across(all_of(numvars$names),
    .fns=list(
      MeanSD=~meansd(.x,
                     add_n = TRUE),
      MedianQuart=median_quart),
    .names = "{.col}: {.fn}")) |>
  pivot_longer(
    cols = everything(),
    names_to = c("Variable",".value"),
    names_sep=":") |>
  flextable() |> 
  set_table_properties(width=1, layout="autofit")
```

## all variables / 2 function / subgroups

<br>

```{r}
# with/without function arguments
clean_data_penguins |> 
  group_by(species) |> 
  summarize(across(
    all_of(numvars$names),
    .fns=list(
      MeanSD=~meansd(.x,
                     add_n = TRUE),
      MedianQuart=median_quart)),
    .groups="drop") |> 
  flextable() |> 
  set_table_properties(width=1, layout="autofit")
```

### variables in rows, groups in columns

<br>

```{r}
clean_data_penguins |> 
  group_by(species) |> 
  summarize(across(
    all_of(numvars$names),
    .fns=list(
      MeanSD=~meansd(.x,
                     add_n = TRUE),
      MedianQuart=median_quart)),
    .groups="drop") |> 
  pivot_longer(
    cols = -species,
    names_to = c("Variable",".value"),
    names_pattern = "^(.+_.+)_(.+)$") |>
  pivot_wider(names_from = species,
              values_from = starts_with("M"),
              names_glue = "{.value}\n{species}",
              names_vary="slowest") |>#paart die Spezies zusammen
  flextable() |> 
  separate_header(split="[:_]") |>
  set_table_properties(width=1, layout="autofit")
```

### with/without function arguments

<br>

```{r}
result_long <- 
  clean_data_penguins |> 
  group_by(species) |> 
  summarize(across(
    all_of(numvars$names),
    .fns=list(
      MeanSD=~meansd(.x,
                     add_n = TRUE),
      MedianQuart=median_quart))) |> 
  pivot_longer(cols = -c(species),
               names_to = c('Variable','.value'),
               names_pattern = "^(.+_.+)_(.+)$",
               values_to = 'Value')
result_long |> 
  flextable() |> 
  merge_v(j=1) |> 
  set_table_properties(width=1, layout="autofit")

cat ("<br>\n\n")

result <- 
  result_long |> 
  pivot_wider(names_from=species,
              names_sep=" ",
              values_from=c(MeanSD, MedianQuart))
result |> 
  flextable() |> 
  separate_header(split="[ ]") |> 
  set_table_properties(width=1, layout="autofit")|>
  theme_booktabs() |>
  align(align = "center", part = "all")
```

## all variables / 2 function / 2 subgroups

```{r}
clean_data_penguins |> 
  group_by(species, sex) |> 
  summarize(across(
    all_of(numvars$names),
    .fns=list(
      MeanSD=~meansd(.x,
                     add_n = TRUE),
      MedianQuart=median_quart)),
    .groups="drop") |> 
  pivot_longer(
    cols = -c(species,sex),
    names_to = "Variable",
    values_to = "MeanSD")|>
  pivot_wider(names_from = "Variable",
              values_from = MeanSD) |>
  flextable() |>
  set_table_properties(width=1, layout="autofit")
```
